---
---

## Exercises

### Exercise 1

Regress the number of hours worked in a week (`WKHP`) against age (`AGEP`). Add a second-order interaction ($$X^2$$) of age to allow for
possible curvature in the relationship. Then, plot the predicted values over the data
to verify the coefficients indicate a downward quadratic relationship.

### Exercise 2

Controlling for a person's educational attainment, fit a linear model that
addresses the question of wage disparity between men and women in the U.S.
workforce. Compare the model to one where sex is not considered.

Add another predictor from the `pums` data frame and evaluate the goodness-of-fit as other predictors are added to the model.

### Exercise 3

Write down the formula for a random intercepts model for earned wages with a
fixed effect of sex and a random effect of educational attainment.

### Exercise 4

Set up a generalized mixed effects model on whether a person attained an advanced
degree (Master's or Doctorate). Include sex and age as fixed effects, and include
a random intercept according to marital status.

===

## Solutions

```{r, title = "Solution 1"}
fit <- lm(
  WKHP ~ AGEP + I(AGEP^2),
  pums)

ggplot(pums,
  aes(x = AGEP, y = WKHP)) +
  geom_point(shape = 'o') +
  geom_line(aes(y = predict(fit)))
```

```{r, title = "Solution 2"}
fit.full <- lm(WAGP ~ SCHL + SEX, pums)
summary(fit.full)
```

```{r, title = "Solution 2"}
fit.schl<- lm(WAGP ~ SCHL, pums)
summary(fit.schl)
```

We can use the `anova` function to compare the two models. 

```{r, title = "Solution 2"}
anova(fit.full, fit.schl, test='Chisq')
```

The two models are significantly different based on the residual sum of squares, a strong indication that sex explains a significant amount of the variation in wages.

We can also add an additional predictor, such as weekly hours worked, `WKHP` to the model. 

```{r, title = "Solution 2"}
fit.wkhp <- lm(WAGP ~ SCHL + SEX + WKHP, data = pums) 
summary(fit.wkhp)
```

Adding `WKHP` to the model increases the $$R^2$$ from `0.19` to `.34`.

```{r, title = "Solution 3"}
WAGP ~ (1 | SCHL) + SEX
```

```{r, title = "Solution 4"}
df <- pums
levels(df$SCHL) <- c(0, 0, 0, 0, 1, 1) 
```

First, we have to manipulate the `pums` data from so that the levels of education are binomial, making education levels of Master's or Doctorate Degrees as 1 and all others as 0. This edited data frame is stored as `df`.

```{r, title = "Solution 4"}
fit <- glmer(
  SCHL ~ (1 | MAR) + AGEP + SEX,
  family = binomial,
  data = df)

summary(fit)
```


