---
---

## Exercises

### Exercise 1

Regress `WKHP` against `AGEP`, adding a second-order interaction to allow for
possible curvature in the relationship. Plot the predicted values over the data
to verify the coefficients indicate a downward quadratic relationship.

### Exercise 2

Controlling for a person's educational attainment, fit a linear model that
addresses the question of wage disparity between men and women in the U.S.
workforce. What other predictor from the `pums` data frame would increase the
goodness-of-fit of the "control" model, before SEX is considered?

### Exercise 3

Set up a generalized mixed effects model on whether a person attained an advanced
degree (Master's or Doctorate). Include sex and age as fixed effects, and include
a random intercept according to occupation.

### Exercise 4

Write down the formula for a random intercepts model for earned wages with a
fixed effect of sex and a random effect of educational attainment.

===

## Solutions

```{r, title = "Solution 1"}
fit <- lm(
  WKHP ~ AGEP + I(AGEP^2),
  pums)

ggplot(pums,
  aes(x = AGEP, y = WKHP)) +
  geom_point(shape = 'o') +
  geom_line(aes(y = predict(fit)))
```

```{r, title = "Solution 2"}
fit <- lm(WAGP ~ SCHL, pums)
summary(fit)
```

The "baseline" model includes educational attainment, when compared to a model
with sex as a second predictor, there is a strong indication of signifigance.

```{r, title = "Solution 2"}
anova(fit, update(fit, WAGP ~ SCHL + SEX))
```

Adding WKHP to the baseline increases the $$R^2$$ to `.32` from `.14` with just
`SCHL`. There is no way to include the possibility that hours worked also
dependends on sex, giving a direct and indirect path to the gender wage gap, in
a linear model.

```{r, title = "Solution 2"}
summary(update(fit, WAGP ~ SCHL + WKHP))
```

```{r, title = "Solution 3"}
df <- pums
levels(df$SCHL) <- c(0, 0, 0, 0, 1, 1)
fit <- glmer(
  SCHL ~ (1 | OCCP) + AGEP,
  family = binomial,
  data = df)
anova(fit, update(fit, . ~ . + SEX), test = 'Chisq')
```

```{r, title = "Solution 4"}
WAGP ~ (1 | SCHL) + SEX
```
