---
---

## Formula Notation

The basic function for fitting a regression in R is the `lm()` function,
standing for linear model.

```{r, handout = 0}
fit <- lm(
  formula = WAGP ~ SCHL,
  data = person)
```

The `lm()` function takes a formula argument and a data argument, and computes
the best fitting linear model (i.e. determines coefficients that [minimize the
sum of squared residuals].
{:.notes}

[minimize the sum of squared residuals]: https://en.wikipedia.org/wiki/Linear_least_squares_(mathematics)

===

Data visualizations can help confirm intuition, but only with very simple
models, typically with just one or two "fixed effects".

```{r}
library(ggplot2)

ggplot(person,
  aes(x = SCHL, y = WAGP)) +
  geom_boxplot()
```

===

### Function, Formula, and Data

The developers of R created an approach to statistical analysis that is both
concise and flexible from the users perspective, while remaining precise and
well-specified for a particular fitting procedure.
{:.notes}

The researcher contemplating a new regression analysis faces three initial
challenges:

1. Choose the function that implements a suitable fitting procedure.
1. Write down the regression model using a formula expression combining
variable names with algebra-like operators.
1. Build a data frame with columns matching these variables that have suitable
data types for the chosen fitting procedure.

The formula expressions are usually very concise, in part because the function
chosen to implement the fitting procedure extracts much necessary information
from the provided data frame. For example, whether a variable is a factor or
numeric is not specified in the formula because it is specified in the data.
{:.notes}

===

## Formula Mini-language(s)
{:style="text-transform: none;"}

- "base R" function `lm()`
- "base R" function `glm()`
- [lme4](){:.rlib} function `lmer()`
- [lme4](){:.rlib} function `glmer()`

Across different packages in R, the formula "mini-language" has evolved to allow
complex model specification. Each package adds syntax to the formula or new
arguments to the fitting function. The `lm` function admits only the simplest
kinds of expressions, while the [lme4](){:.rlib} packagehas evolved this
"mini-language" to allow specification of hierarchichal models.
{:.notes}
