---
---

## Linear Mixed Models

The [lme4](){:.rlib} package expands the formula "mini-language" to allow
descriptions of "random effects" into the linear model. These are called mixed models because they contain both fixed and random effects. 

- normal predictors are "fixed effects"
- `(...|...)` expressions describe "random effects"

In the context of this package, variables
added to the right of the `~` in the usual way are "fixed effects"---they
consume a well-defined number of degrees of freedom. Variables added within
`(...|...)` are "random effects".
{:.notes}

===

Models with random effects should be understood as specifying multiple,
overlapping probability statements about the observations. 

For example, we can to fit the response variable log(WAGP) with the predictors of occupation (`OCCP`) and education (`SCHL`), where `OCCP` is a random effect and `SCHL` is a fixed effect. As assumed by linear models, the response variable, log(WAGP), is normally distributed, with a mean of \mu_i and standard deviation of \sigma_0^2. In addition, the random effect, `OCCP` varies and has a normal distribution, with mean of 0 and standard deviation of \Sigma_1. The fixed effect, `SCHL` does not vary, and thus is a *fixed*.
{..notes}

$$
\begin{align}
&\log(WAGP_i) \sim Normal(\mu_i, \sigma_0^2) \\
&\mu_i = \beta_0 + \boldsymbol{\beta_1}[OCCP_i] + \beta_2 [SCHL_i] \\
&\boldsymbol{\beta_1} \sim Normal(0, \boldsymbol{\Sigma_1}) \\
&\end{align}
$$


===

Mixed models can be fit with "random intercepts" and "random slopes". With a random intercept, each level within the random effect can have a different intercept. With a random slope, each level can have a different slope. 

In `lme4` random effects are shown in the formula inside parentheses, with variables to the right of a "|" to fit random intercept, and to the left to fit a random slope. 

| Formula                  | Description of random effect |
|--------------------------|-------------|
| `y ~ (1 | b) + a`        | random intercept for each level in `b` |
| `y ~ a + (a | b)`        | random intercept and slope w.r.t. `a` for each level in `b` |

===

### Random Intercept

The `lmer` function fits linear models with the [lme4](){:.rlib} extensions to
the `lm` formula syntax.


```{r, message = FALSE, handout = 0}
library(lme4)
fit <- lmer(
  log(WAGP) ~ (1|OCCP) + SCHL,
  data = pums)
```

===

```{r lmer_ri_summary}
summary(fit)
```

The familiar assessment of model residuals is absent from the summary due to the
lack of a widely accepted measure of null and residual deviance. The notions of
model saturation, degrees of freedom, and independence of observations have all
crossed onto thin ice.
{:.notes}

In a `lm` or `glm` fit, each response is conditionally independent, given it's
predictors and the model coefficients. Each observation corresponds to it's own
probability statement. In a model with random effects, each response is no
longer conditionally independent, given it's predictors and model coefficients.
{:.notes}

===

### Random Slope

Adding a numeric variable on the left of a grouping specified with `(...|...)`
produces a "random slope" model. Here, separate coefficients for hours worked
are allowed for each level of education.

```{r, handout = 0}
fit <- lmer(
  log(WAGP) ~ (WKHP | SCHL),
  data = pums)
```

===

The warning that the procedure `failed to converge` is worth paying attention
to. One way to fix this is to change the numerical optimizer used for model fitting
that [lme4](){:.rlib} is used by default. In this example, we change the optimizer to a "BOBYQA" optimizer, which is slower than the default. 

```{r, handout = 0}
fit <- lmer(
  log(WAGP) ~ (WKHP | SCHL),
  data = pums, 
  control = lmerControl(optimizer = "bobyqa"))
```

```{r}
summary(fit)
```

===

The `predict` function extracts model predictions from a fitted model object
(i.e. the output of `lm`, `glm`, `lmer`, and `glmer`), providing one easy
way to visualize the effect of estimated coefficients. We can plot the extracted coefficients using `ggplot`.

```{r, handout = 0}
ggplot(pums,
  aes(x = WKHP, y = log(WAGP), color = SCHL)) +
  geom_point() +
  geom_line(aes(y = predict(fit))) +
  labs(title = 'Random intercept and slope with lmer')
```

===

### Generalized Mixed Models

The `glmer` function adds upon `lmer` the option to specify the same group of
exponential family distributions as `glm` adds to `lm`. The same `family` argument is used in `glmer` to specify the distribution. 